---
title: "IS6481 | Lab 6 -- Supervised Learning Responses"
output: 
  html_document: 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# options(width = 120)
library(ggplot2)
library(fpc)
library(tidyverse)
library(reshape2)
library(rpart)
library(rpart.plot)


mm <- src_mysql(
  host='is6481-mysql.cffidl3kiu88.us-east-1.rds.amazonaws.com',
  port=3306,
  user='root',
  password='test1234',
  dbname='supervised_learning'
)
```

```{r data_acquisition,echo=FALSE}
d_weather <- tbl(mm,'weather') %>% collect()
```

```{r data_cleaning,echo=FALSE}
d_weather_remove <- d_weather %>% select(-Location)
d_weather_date <- d_weather_remove %>% mutate(Date=as.Date(Date,format='%m/%d/%y'))
```


## Data cleaning

Answer the following question.

* *List two different approaches for dealing with attributes that have missing (i.e. NA) values.*

If it was a requirement that all explanatory attributes were numeric, how might you handle:

1. *Date*
2. *WindGustDir*


## Data Exploration
Answer the following questions regarding the following two charts seen in the lab.
```{r explore_viz,echo=FALSE}
ggplot(d_weather, aes(x=WindGustDir)) + geom_bar() + coord_flip()
ggplot(d_weather, aes(x=as.numeric(MinTemp))) + geom_density()

```

* *Explain each of the plots and what they could mean*
* *What do these plots tell us about how accurate the model might be?*

## Model Training and Accuracy

```{r data model,echo=FALSE}
decision_tree_model <- rpart(RainTomorrow ~ MinTemp + MaxTemp + Rainfall + Evaporation + Sunshine + WindGustDir + WindGustSpeed,data=d_weather_date)
```

Answer the following regarding model accuracy.

* *If we care most about predicting when it rains (i.e. rain has the biggest impact on us), which metrics related to the confusion matrix are important to consider? Meaning, do we care more or less about the overall prediction accuracy and is there a subset of the prediction accuracy we care more about? See: [https://en.wikipedia.org/wiki/Confusion_matrix](https://en.wikipedia.org/wiki/Confusion_matrix)*
* *Assume we care most about predicting when it rains.  Calculate the metrics related to the confusion matrix that are most important.*

## Prediction

Answer the following regarding the chart below.

* *What does the model predict if Sunshine = 7.3 and MinTemp = 15?*
* *What does the model predict if Sunshine = 8, WindGust = E?*


```{r plot_tree,echo=FALSE}
prp(decision_tree_model, extra = 1) 
``` 